---
title: 'Developing Data Products: Premier League 2019/20 Stadiums'
author: "Park"
date: "9/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Executive Summary

This project attempts to create an interactive map which pins the locations of each of the Premier League 2019/20 football team's home stadium. The data for the locations of the stadiums comes from this [Wikipedia page](https://en.wikipedia.org/wiki/List_of_Premier_League_stadiums).

## Premier League 2019/20 Stadiums

### Stadiums Data
This project starts from scratch. Thus, the first thing to do is to create the *stadium* data frame which consists of the following columns:  
1) Club -- Name of the club which owns the stadium  
2) Name -- Name of the stadium  
3) lat -- The latitude coordinate of each stadium round to 6 decimal places  
4) lng -- The longitude coordinate of each stadium round to 6 decimal places  
5) Capacity -- The number of seats available in the stadium  
6) Position -- The finishing position of the 2019/20 season  
7) Color -- Classifies all teams according to European football qualification and relegation to Championship     
8) Status -- Which 2020/21 European competition the team qualifies to  

Please note that the data frame follows the alphabetical order of the name of the football club.

``` {r create, echo = TRUE}
stadium <- data.frame(
    Club = c("Arsenal", "Aston Villa", "Bournemouth", "Brighton & Hove Albion", 
             "Burnley", "Chelsea", "Crystal Palace", "Everton", "Leicester City", 
             "Liverpool", "Manchester City", "Manchester United", "Newcastle United", 
             "Norwich City", "Sheffield United", "Southampton", "Tottenham Hotspur", 
             "Watford", "West Ham Umited", "Wolverhampton Wanderers"),
    Name = c("Emirates Stadium", "Villa Park", "Dean Court", "Falmer Stadium", 
             "Turf Moor", "Stamford Bridge", "Selhurst Park", "Goodison Park", 
             "King Power Stadium", "Anfield", "City of Manchester Stadium", 
             "Old Trafford", "St James' Park", "Carrow Road", "Bramall Lane", 
             "St Mary's Stadium", "Tottenham Hotspur Stadium", "Vicarage Road", 
             "London Stadium", "Molineux Stadium"),
    lat = c(51.555, 52.509167, 50.735278, 50.861822, 53.789167, 51.481667, 
             51.398333, 53.438889, 52.620278, 53.430833, 53.483056, 53.463056, 
             54.975556, 52.622222, 53.370278, 50.905833, 51.604722, 51.65, 
             51.538611, 52.590278),
    lng = c(-0.108611, -1.884722, -1.838333, -0.083278, -2.230278, -0.191111, 
             -0.085556, -2.966389, -1.142222, -2.960833, -2.200278, -2.291389, 
             -1.621667, 1.309167, -1.470833, -1.391111, -0.066389, -0.401667, 
             -0.016389, -2.130278),
    Capacity = c(60704, 42095, 11364, 30750, 21944, 40834, 25486, 39414, 32361, 
                 53394, 55017, 75643, 52305, 27359, 32125, 32384, 62214, 22200, 
                 60000, 32050),
    Position = c(8, 17, 18, 15, 10, 4, 14, 12, 5, 1, 2, 3, 13, 20, 9, 11, 6, 19, 
                 16, 7),
    Color = c("orange", "grey", "red", "grey", "grey", "green", "grey", "grey", "orange", "blue", "green", "green", "grey", "red", "grey", "grey", "orange", "red", "grey", "grey")
)

# add the status of each team for the 2020/21 season
stadium$Status <- ifelse(stadium$Color == "blue","PL Champion and Champions League 2020/21",
    ifelse(stadium$Color == "green", "Champions League 2020/21", 
    ifelse(stadium$Color == "orange", "Europa League 2020/21", 
    ifelse(stadium$Color == "red", "Relegated to Championship", "None"))))
```

### Stadiums Map

Once the data is ready, the next step is to create an interactive map. This is done using the *leaflet* package in R.

Apart from displaying the locations of each stadium, the size of the markers illustrates another dimension of the data. That is, the **capacity** of the stadiums determines how large the markers are. In addition, the **colors** of the markers feature each club's qualification in the upcoming 2020/21 season. There are 5 colors as followed:

- Blue: Premier League 2019/20 champion and qualifies to the 2020/21 Champions League
- Green: Qualify to the 2020/21 Champions League
- Orange: Qualify to the 2020/21 Europa League
- Grey: No European football qualification
- Red: Relegated to the 2020/21 Championship League

Please note that in addition to the finishing position in the 2019/20 season, the title holders and runner-ups (if any) for 2019/20 domestic and European competitions may also qualify to the next season European competitions.

``` {r map, echo = TRUE}
library(leaflet); library(htmltools) #for br()
stadium %>% 
    leaflet() %>% 
    addTiles() %>% 
    # add markers
    addCircleMarkers(fillColor = stadium$Color, color = stadium$Color, opacity = 0.8, weight = 3, radius = stadium$Capacity/2500, fillOpacity = 0.4, popup = paste(stadium$Club,":", " ", stadium$Name, br(), "Capacity:", " ", stadium$Capacity, sep = ""), clusterOptions = markerClusterOptions()) %>%
    # add legends
    addLegend(labels = c("PL Champion and Champions League 2020/21", "Champions League 2020/21", "Europa League 2020/21", "Relegated"), colors = c("blue", "green", "orange", "red"), position = "topright")
```

If clicked on each of the marker, some information will be shown. This includes the names of the club and the stadium and the stadium capacity.  
  
  
  

